Sophize Markdown and Collaboration Interface

### 2021-10-17 math_library/sophize_math.tex
changes

@@

\section{Sophize's Knowledge Organization Scheme}

\subsection{Sophize's Data Model}
\paragraph{Before we describe the Sophize Markdown language, it will be necessary to briefly describe how Sophize models mathematical entities. The data model is published in JSON schema \cite{sophize_datamodel} and popular package managers such as MVN, npm and PyPI. We describe relevant parts of the concepts used here.}

\paragraph{A \textbf{resource} is an abstract concept inherited by all other top-level concepts like terms, propositions, and arguments. Each resource has a URI and contains fields such as search tags and citations.}

@@

\subsection{Polymath Projects}
In order to better understand the nature of online collaborations, we studied the Polymath projects \cite{polymath_blog}, which are massively collaborative online mathematical projects. Conveniently, the project leaders publically analyzed these projects and suggested technical improvements suitable for such collaborations. In fact, a large portion of the interface is designed specifically to overcome the issues noted by Timothy Gowers \cite{gowers_weblog_2009} and Terence Tao \cite{whats_new_2009}.

>>>>

- The goal is rational truth. Three dimensions are consistency, correspondence and values.

- In its current iteration, Sophize can modekl vast mnajority (almost all)  of mathematics.

@@

\section{Sophize's Knowledge Organization Scheme}
For math logical consistency is the sole criteria for truth..
Thus, as a math library, Sophize models mathematical knowledge using logical arguments.

The various concepts used for orgnizing knowledge logically are described below. These concepts have been modelled in JSON schema \cite{sophize_datamodel} and published for popular programming language such as Java (MVN), Javascript (npm) and Python (PyPI).

@@

\paragraph{While modelling the concepts an abstract concept for ease of implementation. A \textbf{resource} is an abstract concept inherited by all other top-level concepts like terms, propositions, and arguments. Each resource has a URI and contains fields such as search tags and citations.}

The same URI scheme will be used in this paper to refer to various terms, propositions, arguments, etc.

There are a couple more concepts that we will describe at later in the paper.

\subsection{Belief sets}

In Sophize, a proposition is claimed to be true if it is concluded by a valid argument and the premises of the argument are also true.

Of course this means that we need to start with some propositions that are assumed to be true. Traditionally, these are mathematical axioms and are considered to true because they are self-evident. However, there is no criteria to define if a propositions if self-evidently true. In mathematical inquiry, theorems are proven/disproven in one or more multiple theories, each with its own set of axioms. Each theory may have its theoretical advantages and practical uses and there are no criteria to choose one theory over another when trying to assert a proposition's truth/falsity.

Thus, Sophize allows any set of propositions to be considered true without evidence. Such a set of propositions is called a belief set. Thus, the truth-value of a proposition is only defined within a belief set - it doesn't have any concept of universal truth.

The same argument may be utilized by multiple belief sets to compute truth values.

Computing of truth values in a belief set is thus a simple recursive graph algorithm. Sophize computes truth values for all its belief sets.

\subsection{Validity of arguments}
Validity of an argument is, in principle, objectively verifiable. A formally defined logistic systems have a mechanical procedure to verify validity arguments.

However, mathematics only utilizes a tiny portion formally verifiable arguments. Most mathematics is verifies via peer review, some with curated by experts.

\subsection{Tracking Inconsistency}
If a propositions and its negation is proved within a belief set is inconsistent. As proved by the principle of explosion, any propositions can be proved from such a contradiction. Thus making the claim of truth of any proposition practically worthless.

Thus it is important to track and report contradictions in a belief set. Thus Sophize has a semantic notion of negation of a proposition. The premises or conclusion of an argument can be negation of propositions.

When a proposition and its negation are proven, the system reports such contradictions for every belief set.
